---
title: "Getting back to work on the med data"
author: "Christopher Morano"
date: "September 11, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(randomForest)
library(xgboost)
library(caret)
library(lime)
library(extrafont)
library(Amelia) #to visualize missing data
library(printr)
```

```{r}
#Loading the Data:
Aug_med <- read_csv('../data/Med_data_Aug21.csv', col_types=list("Day4"=col_character(), 
                                                                 "Day6"=col_character(),
                                                                 "Day7"=col_character()))
#Let's take the first 21 days of medical results:
Aug_med_day21 <- select(Aug_med, ID:Day21)

#We need to clean up columns Day4, 6, & 7, but this would be easier when I make the data tall.
```

```{r, fig.height=5, fig.width=10}
missmap(Aug_med_day21)
```

Let's save a bunch of versions of this data so we can easily run models on 0 vs. ABC, A vs. B vs. C, A vs. BC, AB vs. C, and 0AB vs. C.  That's five versions I can think of right now.  Let's start with those.

```{r}
#Let's get our five data sets:

#Aug_med_0vsABC
#___________________

Aug_med_0vsABC <- Aug_med_day21
Aug_med_0vsABC[Aug_med_0vsABC$Result2016 %in% c("A", "B", "C"),]$Result2016 <- "ABC"
Aug_med_0vsABC$Result2016 <- factor(Aug_med_0vsABC$Result2016)

#Aug_med_0ABvsC
#___________________

Aug_med_0ABvsC <- Aug_med_day21
Aug_med_0ABvsC[Aug_med_0ABvsC$Result2016 %in% c("0", "A", "B"),]$Result2016 <- "0AB"
Aug_med_0ABvsC$Result2016 <- factor(Aug_med_0ABvsC$Result2016)

#Aug_med_AvsBvsC
#___________________

Aug_med_AvsBvsC <- filter(Aug_med_day21, Result2016 %in% c("A", "B", "C"))
Aug_med_AvsBvsC$Result2016 <- factor(Aug_med_AvsBvsC$Result2016)

#Aug_med_ABvsC
#___________________

Aug_med_ABvsC <- filter(Aug_med_day21, Result2016 %in% c("A", "B", "C"))
Aug_med_ABvsC[Aug_med_ABvsC$Result2016 %in% c("A", "B"),]$Result2016 <- "AB"
Aug_med_ABvsC$Result2016 <- factor(Aug_med_ABvsC$Result2016)

#Aug_med_AvsBC
#___________________

Aug_med_AvsBC <- filter(Aug_med_day21, Result2016 %in% c("A", "B", "C"))
Aug_med_AvsBC[Aug_med_AvsBC$Result2016 %in% c("B", "C"),]$Result2016 <- "BC"
Aug_med_AvsBC$Result2016 <- factor(Aug_med_AvsBC$Result2016)

#table(Aug_med_AvsBC$Result2016)
```

Let's summarize and see how far we can get with just the summary of 21 days.  We can shorten it later, but I'm wondering if having 3 weeks of data makes a big difference (compared to just 1 week).


```{r}
#Tall: 0 vs. ABC
Aug_med_0vsABC.tall.21 <- gather(Aug_med_0vsABC, day, result, Day1:Day21) %>% 
  filter(!is.na(result)) %>% 
  arrange(ID)

Aug_med_0vsABC.tall.7 <- gather(select(Aug_med_0vsABC, ID:Day7), day, result, Day1:Day7) %>% 
  filter(!is.na(result)) %>% 
  arrange(ID)
```

```{r}
#Tall: 0AB vs. C
Aug_med_0ABvsC.tall.21 <- gather(Aug_med_0ABvsC, day, result, Day1:Day21) %>% 
  filter(!is.na(result)) %>% 
  arrange(ID)

Aug_med_0ABvsC.tall.7 <- gather(select(Aug_med_0ABvsC, ID:Day7), day, result, Day1:Day7) %>% 
  filter(!is.na(result)) %>% 
  arrange(ID)
```

```{r}
#Tall: 0AB vs. C
Aug_med_AvsBvsC.tall.21 <- gather(Aug_med_AvsBvsC, day, result, Day1:Day21) %>% 
  filter(!is.na(result)) %>% 
  arrange(ID)

Aug_med_AvsBvsC.tall.7 <- gather(select(Aug_med_AvsBvsC, ID:Day7), day, result, Day1:Day7) %>% 
  filter(!is.na(result)) %>% 
  arrange(ID)
```

```{r}
#Tall: AB vs. C
Aug_med_ABvsC.tall.21 <- gather(Aug_med_ABvsC, day, result, Day1:Day21) %>% 
  filter(!is.na(result)) %>% 
  arrange(ID)

Aug_med_ABvsC.tall.7 <- gather(select(Aug_med_ABvsC, ID:Day7), day, result, Day1:Day7) %>% 
  filter(!is.na(result)) %>% 
  arrange(ID)
```

```{r}
#Tall: A vs. BC
Aug_med_AvsBC.tall.21 <- gather(Aug_med_AvsBC, day, result, Day1:Day21) %>% 
  filter(!is.na(result)) %>% 
  arrange(ID)

Aug_med_AvsBC.tall.7 <- gather(select(Aug_med_AvsBC, ID:Day7), day, result, Day1:Day7) %>% 
  filter(!is.na(result)) %>% 
  arrange(ID)
```

```{r}
#cleaning up all of the data!!!:
#first: convert `day` column to ordered factors:
Aug_med_0vsABC.tall.21$day <- ordered(Aug_med_0vsABC.tall.21$day)
Aug_med_0ABvsC.tall.21$day <- ordered(Aug_med_0ABvsC.tall.21$day)
Aug_med_AvsBvsC.tall.21$day <- ordered(Aug_med_AvsBvsC.tall.21$day)
Aug_med_AvsBC.tall.21$day <- ordered(Aug_med_AvsBC.tall.21$day)
Aug_med_ABvsC.tall.21$day <- ordered(Aug_med_ABvsC.tall.21$day)
Aug_med_0vsABC.tall.7$day <- ordered(Aug_med_0vsABC.tall.7$day)
Aug_med_0ABvsC.tall.7$day <- ordered(Aug_med_0ABvsC.tall.7$day)
Aug_med_AvsBvsC.tall.7$day <- ordered(Aug_med_AvsBvsC.tall.7$day)
Aug_med_AvsBC.tall.7$day <- ordered(Aug_med_AvsBC.tall.7$day)
Aug_med_ABvsC.tall.7$day <- ordered(Aug_med_ABvsC.tall.7$day)

#changing the values of `<10` and `>30000`:
Aug_med_0vsABC.tall.21$result[Aug_med_0vsABC.tall.21$result == '>30000'] <- '30001'
Aug_med_0ABvsC.tall.21$result[Aug_med_0ABvsC.tall.21$result == '>30000'] <- '30001'
Aug_med_AvsBvsC.tall.21$result[Aug_med_AvsBvsC.tall.21$result == '>30000'] <- '30001'
Aug_med_AvsBC.tall.21$result[Aug_med_AvsBC.tall.21$result == '>30000'] <- '30001'
Aug_med_ABvsC.tall.21$result[Aug_med_ABvsC.tall.21$result == '>30000'] <- '30001'
Aug_med_0vsABC.tall.7$result[Aug_med_0vsABC.tall.7$result == '>30000'] <- '30001'
Aug_med_0ABvsC.tall.7$result[Aug_med_0ABvsC.tall.7$result == '>30000'] <- '30001'
Aug_med_AvsBvsC.tall.7$result[Aug_med_AvsBvsC.tall.7$result == '>30000'] <- '30001'
Aug_med_AvsBC.tall.7$result[Aug_med_AvsBC.tall.7$result == '>30000'] <- '30001'
Aug_med_ABvsC.tall.7$result[Aug_med_ABvsC.tall.7$result == '>30000'] <- '30001'

set.seed(1313)
random10 <- round(runif(10, 1, 9))
random12 <- round(runif(12, 1, 9))

Aug_med_0vsABC.tall.21$result[Aug_med_0vsABC.tall.21$result == '<10'] <- random12
Aug_med_0ABvsC.tall.21$result[Aug_med_0ABvsC.tall.21$result == '<10'] <- random12
Aug_med_AvsBvsC.tall.21$result[Aug_med_AvsBvsC.tall.21$result == '<10'] <- random10
Aug_med_AvsBC.tall.21$result[Aug_med_AvsBC.tall.21$result == '<10'] <- random10
Aug_med_ABvsC.tall.21$result[Aug_med_ABvsC.tall.21$result == '<10'] <- random10
Aug_med_0vsABC.tall.7$result[Aug_med_0vsABC.tall.7$result == '<10'] <- random12
Aug_med_0ABvsC.tall.7$result[Aug_med_0ABvsC.tall.7$result == '<10'] <- random12
Aug_med_AvsBvsC.tall.7$result[Aug_med_AvsBvsC.tall.7$result == '<10'] <- random10
Aug_med_AvsBC.tall.7$result[Aug_med_AvsBC.tall.7$result == '<10'] <- random10
Aug_med_ABvsC.tall.7$result[Aug_med_ABvsC.tall.7$result == '<10'] <- random10

#converting them all to numeric:
Aug_med_0vsABC.tall.21$result <- as.numeric(Aug_med_0vsABC.tall.21$result)
Aug_med_0ABvsC.tall.21$result <- as.numeric(Aug_med_0ABvsC.tall.21$result)
Aug_med_AvsBvsC.tall.21$result <- as.numeric(Aug_med_AvsBvsC.tall.21$result)
Aug_med_AvsBC.tall.21$result <- as.numeric(Aug_med_AvsBC.tall.21$result)
Aug_med_ABvsC.tall.21$result <- as.numeric(Aug_med_ABvsC.tall.21$result)
Aug_med_0vsABC.tall.7$result <- as.numeric(Aug_med_0vsABC.tall.7$result)
Aug_med_0ABvsC.tall.7$result <- as.numeric(Aug_med_0ABvsC.tall.7$result)
Aug_med_AvsBvsC.tall.7$result <- as.numeric(Aug_med_AvsBvsC.tall.7$result)
Aug_med_AvsBC.tall.7$result <- as.numeric(Aug_med_AvsBC.tall.7$result)
Aug_med_ABvsC.tall.7$result <- as.numeric(Aug_med_ABvsC.tall.7$result)

Aug_med_0vsABC.tall.21
#I could have done this with a loop...  Oh well...
```

```{r}
#3-week data:
#Aug_med_0vsABC.sum.21
Aug_med_0vsABC.sum.21 <- Aug_med_0vsABC.tall.21 %>% 
  group_by(ID) %>% 
  summarize(Gender=max(Gender),
            Age=max(Age),
            BMI=max(BMI),
            AsaScore=max(AsaScore),
            TexturePancreas=max(TexturePancreas),
            MPD_Diameter=max(MPD_Diameter),
            DGE_YesOrNot=max(DGE_YesOrNot),
            POPF=max(POPF),
            Bleeding=max(Bleeding),
            Reoperation=max(Reoperation),
            EndoscopicReintervention=max(EndoscopicReintervention),
            RadiologicalReintervention=max(RadiologicalReintervention),
            OrganFailure=max(OrganFailure),
            InHospitalMortality=max(InHospitalMortality),
            HospitalReadmission=max(HospitalReadmission),
            HospitalStay=max(HospitalStay),
            POD_DrainRemoval=max(POD_DrainRemoval),
            num_of_results=n(),
            min_result=min(result),
            mean_result=mean(result),
            max_result=max(result),
            var_result=var(result),
            last_result=day[num_of_results],
            diff_of_last_and_max=ifelse(num_of_results==1, NA, max_result-result[num_of_results]),
            diff_of_last_and_min=ifelse(num_of_results==1, NA, result[num_of_results]-min_result),
            Result_2016=factor(names(which.max(table(Result2016)))))

#Aug_med_0ABvsC.sum.21
Aug_med_0ABvsC.sum.21 <- Aug_med_0ABvsC.tall.21 %>% 
  group_by(ID) %>% 
  summarize(Gender=max(Gender),
            Age=max(Age),
            BMI=max(BMI),
            AsaScore=max(AsaScore),
            TexturePancreas=max(TexturePancreas),
            MPD_Diameter=max(MPD_Diameter),
            DGE_YesOrNot=max(DGE_YesOrNot),
            POPF=max(POPF),
            Bleeding=max(Bleeding),
            Reoperation=max(Reoperation),
            EndoscopicReintervention=max(EndoscopicReintervention),
            RadiologicalReintervention=max(RadiologicalReintervention),
            OrganFailure=max(OrganFailure),
            InHospitalMortality=max(InHospitalMortality),
            HospitalReadmission=max(HospitalReadmission),
            HospitalStay=max(HospitalStay),
            POD_DrainRemoval=max(POD_DrainRemoval),
            num_of_results=n(),
            min_result=min(result),
            mean_result=mean(result),
            max_result=max(result),
            var_result=var(result),
            last_result=day[num_of_results],
            diff_of_last_and_max=ifelse(num_of_results==1, NA, max_result-result[num_of_results]),
            diff_of_last_and_min=ifelse(num_of_results==1, NA, result[num_of_results]-min_result),
            Result_2016=factor(names(which.max(table(Result2016)))))

#Aug_med_AvsBvsC.sum.21
Aug_med_AvsBvsC.sum.21 <- Aug_med_AvsBvsC.tall.21 %>% 
  group_by(ID) %>% 
  summarize(Gender=max(Gender),
            Age=max(Age),
            BMI=max(BMI),
            AsaScore=max(AsaScore),
            TexturePancreas=max(TexturePancreas),
            MPD_Diameter=max(MPD_Diameter),
            DGE_YesOrNot=max(DGE_YesOrNot),
            POPF=max(POPF),
            Bleeding=max(Bleeding),
            Reoperation=max(Reoperation),
            EndoscopicReintervention=max(EndoscopicReintervention),
            RadiologicalReintervention=max(RadiologicalReintervention),
            OrganFailure=max(OrganFailure),
            InHospitalMortality=max(InHospitalMortality),
            HospitalReadmission=max(HospitalReadmission),
            HospitalStay=max(HospitalStay),
            POD_DrainRemoval=max(POD_DrainRemoval),
            num_of_results=n(),
            min_result=min(result),
            mean_result=mean(result),
            max_result=max(result),
            var_result=var(result),
            last_result=day[num_of_results],
            diff_of_last_and_max=ifelse(num_of_results==1, NA, max_result-result[num_of_results]),
            diff_of_last_and_min=ifelse(num_of_results==1, NA, result[num_of_results]-min_result),
            Result_2016=factor(names(which.max(table(Result2016)))))

#Aug_med_ABvsC.sum.21
Aug_med_ABvsC.sum.21 <- Aug_med_ABvsC.tall.21 %>% 
  group_by(ID) %>% 
  summarize(Gender=max(Gender),
            Age=max(Age),
            BMI=max(BMI),
            AsaScore=max(AsaScore),
            TexturePancreas=max(TexturePancreas),
            MPD_Diameter=max(MPD_Diameter),
            DGE_YesOrNot=max(DGE_YesOrNot),
            POPF=max(POPF),
            Bleeding=max(Bleeding),
            Reoperation=max(Reoperation),
            EndoscopicReintervention=max(EndoscopicReintervention),
            RadiologicalReintervention=max(RadiologicalReintervention),
            OrganFailure=max(OrganFailure),
            InHospitalMortality=max(InHospitalMortality),
            HospitalReadmission=max(HospitalReadmission),
            HospitalStay=max(HospitalStay),
            POD_DrainRemoval=max(POD_DrainRemoval),
            num_of_results=n(),
            min_result=min(result),
            mean_result=mean(result),
            max_result=max(result),
            var_result=var(result),
            last_result=day[num_of_results],
            diff_of_last_and_max=ifelse(num_of_results==1, NA, max_result-result[num_of_results]),
            diff_of_last_and_min=ifelse(num_of_results==1, NA, result[num_of_results]-min_result),
            Result_2016=factor(names(which.max(table(Result2016)))))

#Aug_med_AvsBC.sum.21
Aug_med_AvsBC.sum.21 <- Aug_med_AvsBC.tall.21 %>% 
  group_by(ID) %>% 
  summarize(Gender=max(Gender),
            Age=max(Age),
            BMI=max(BMI),
            AsaScore=max(AsaScore),
            TexturePancreas=max(TexturePancreas),
            MPD_Diameter=max(MPD_Diameter),
            DGE_YesOrNot=max(DGE_YesOrNot),
            POPF=max(POPF),
            Bleeding=max(Bleeding),
            Reoperation=max(Reoperation),
            EndoscopicReintervention=max(EndoscopicReintervention),
            RadiologicalReintervention=max(RadiologicalReintervention),
            OrganFailure=max(OrganFailure),
            InHospitalMortality=max(InHospitalMortality),
            HospitalReadmission=max(HospitalReadmission),
            HospitalStay=max(HospitalStay),
            POD_DrainRemoval=max(POD_DrainRemoval),
            num_of_results=n(),
            min_result=min(result),
            mean_result=mean(result),
            max_result=max(result),
            var_result=var(result),
            last_result=day[num_of_results],
            diff_of_last_and_max=ifelse(num_of_results==1, NA, max_result-result[num_of_results]),
            diff_of_last_and_min=ifelse(num_of_results==1, NA, result[num_of_results]-min_result),
            Result_2016=factor(names(which.max(table(Result2016)))))

```

```{r}
#1-week data:
#Aug_med_0vsABC.sum.7
Aug_med_0vsABC.sum.7 <- Aug_med_0vsABC.tall.7 %>% 
  group_by(ID) %>% 
  summarize(Gender=max(Gender),
            Age=max(Age),
            BMI=max(BMI),
            AsaScore=max(AsaScore),
            TexturePancreas=max(TexturePancreas),
            MPD_Diameter=max(MPD_Diameter),
            DGE_YesOrNot=max(DGE_YesOrNot),
            POPF=max(POPF),
            Bleeding=max(Bleeding),
            Reoperation=max(Reoperation),
            EndoscopicReintervention=max(EndoscopicReintervention),
            RadiologicalReintervention=max(RadiologicalReintervention),
            OrganFailure=max(OrganFailure),
            InHospitalMortality=max(InHospitalMortality),
            HospitalReadmission=max(HospitalReadmission),
            HospitalStay=max(HospitalStay),
            POD_DrainRemoval=max(POD_DrainRemoval),
            num_of_results=n(),
            min_result=min(result),
            mean_result=mean(result),
            max_result=max(result),
            var_result=var(result),
            last_result=day[num_of_results],
            diff_of_last_and_max=ifelse(num_of_results==1, NA, max_result-result[num_of_results]),
            diff_of_last_and_min=ifelse(num_of_results==1, NA, result[num_of_results]-min_result),
            Result_2016=factor(names(which.max(table(Result2016)))))

#Aug_med_0ABvsC.sum.7
Aug_med_0ABvsC.sum.7 <- Aug_med_0ABvsC.tall.7 %>% 
  group_by(ID) %>% 
  summarize(Gender=max(Gender),
            Age=max(Age),
            BMI=max(BMI),
            AsaScore=max(AsaScore),
            TexturePancreas=max(TexturePancreas),
            MPD_Diameter=max(MPD_Diameter),
            DGE_YesOrNot=max(DGE_YesOrNot),
            POPF=max(POPF),
            Bleeding=max(Bleeding),
            Reoperation=max(Reoperation),
            EndoscopicReintervention=max(EndoscopicReintervention),
            RadiologicalReintervention=max(RadiologicalReintervention),
            OrganFailure=max(OrganFailure),
            InHospitalMortality=max(InHospitalMortality),
            HospitalReadmission=max(HospitalReadmission),
            HospitalStay=max(HospitalStay),
            POD_DrainRemoval=max(POD_DrainRemoval),
            num_of_results=n(),
            min_result=min(result),
            mean_result=mean(result),
            max_result=max(result),
            var_result=var(result),
            last_result=day[num_of_results],
            diff_of_last_and_max=ifelse(num_of_results==1, NA, max_result-result[num_of_results]),
            diff_of_last_and_min=ifelse(num_of_results==1, NA, result[num_of_results]-min_result),
            Result_2016=factor(names(which.max(table(Result2016)))))

#Aug_med_AvsBvsC.sum.7
Aug_med_AvsBvsC.sum.7 <- Aug_med_AvsBvsC.tall.7 %>% 
  group_by(ID) %>% 
  summarize(Gender=max(Gender),
            Age=max(Age),
            BMI=max(BMI),
            AsaScore=max(AsaScore),
            TexturePancreas=max(TexturePancreas),
            MPD_Diameter=max(MPD_Diameter),
            DGE_YesOrNot=max(DGE_YesOrNot),
            POPF=max(POPF),
            Bleeding=max(Bleeding),
            Reoperation=max(Reoperation),
            EndoscopicReintervention=max(EndoscopicReintervention),
            RadiologicalReintervention=max(RadiologicalReintervention),
            OrganFailure=max(OrganFailure),
            InHospitalMortality=max(InHospitalMortality),
            HospitalReadmission=max(HospitalReadmission),
            HospitalStay=max(HospitalStay),
            POD_DrainRemoval=max(POD_DrainRemoval),
            num_of_results=n(),
            min_result=min(result),
            mean_result=mean(result),
            max_result=max(result),
            var_result=var(result),
            last_result=day[num_of_results],
            diff_of_last_and_max=ifelse(num_of_results==1, NA, max_result-result[num_of_results]),
            diff_of_last_and_min=ifelse(num_of_results==1, NA, result[num_of_results]-min_result),
            Result_2016=factor(names(which.max(table(Result2016)))))

#Aug_med_ABvsC.sum.7
Aug_med_ABvsC.sum.7 <- Aug_med_ABvsC.tall.7 %>% 
  group_by(ID) %>% 
  summarize(Gender=max(Gender),
            Age=max(Age),
            BMI=max(BMI),
            AsaScore=max(AsaScore),
            TexturePancreas=max(TexturePancreas),
            MPD_Diameter=max(MPD_Diameter),
            DGE_YesOrNot=max(DGE_YesOrNot),
            POPF=max(POPF),
            Bleeding=max(Bleeding),
            Reoperation=max(Reoperation),
            EndoscopicReintervention=max(EndoscopicReintervention),
            RadiologicalReintervention=max(RadiologicalReintervention),
            OrganFailure=max(OrganFailure),
            InHospitalMortality=max(InHospitalMortality),
            HospitalReadmission=max(HospitalReadmission),
            HospitalStay=max(HospitalStay),
            POD_DrainRemoval=max(POD_DrainRemoval),
            num_of_results=n(),
            min_result=min(result),
            mean_result=mean(result),
            max_result=max(result),
            var_result=var(result),
            last_result=day[num_of_results],
            diff_of_last_and_max=ifelse(num_of_results==1, NA, max_result-result[num_of_results]),
            diff_of_last_and_min=ifelse(num_of_results==1, NA, result[num_of_results]-min_result),
            Result_2016=factor(names(which.max(table(Result2016)))))

#Aug_med_AvsBC.sum.7
Aug_med_AvsBC.sum.7 <- Aug_med_AvsBC.tall.7 %>% 
  group_by(ID) %>% 
  summarize(Gender=max(Gender),
            Age=max(Age),
            BMI=max(BMI),
            AsaScore=max(AsaScore),
            TexturePancreas=max(TexturePancreas),
            MPD_Diameter=max(MPD_Diameter),
            DGE_YesOrNot=max(DGE_YesOrNot),
            POPF=max(POPF),
            Bleeding=max(Bleeding),
            Reoperation=max(Reoperation),
            EndoscopicReintervention=max(EndoscopicReintervention),
            RadiologicalReintervention=max(RadiologicalReintervention),
            OrganFailure=max(OrganFailure),
            InHospitalMortality=max(InHospitalMortality),
            HospitalReadmission=max(HospitalReadmission),
            HospitalStay=max(HospitalStay),
            POD_DrainRemoval=max(POD_DrainRemoval),
            num_of_results=n(),
            min_result=min(result),
            mean_result=mean(result),
            max_result=max(result),
            var_result=var(result),
            last_result=day[num_of_results],
            diff_of_last_and_max=ifelse(num_of_results==1, NA, max_result-result[num_of_results]),
            diff_of_last_and_min=ifelse(num_of_results==1, NA, result[num_of_results]-min_result),
            Result_2016=factor(names(which.max(table(Result2016)))))

```

Okay!  Let's get to work!

The goal is to see how well we can predict with each data set.  Can we just use the results obtained from aggregating the test results, or do we need other factors?  

Originally, I got great results for 0 vs ABC using a random forest for 1-week's worth of data, using just the test results.  Let's double-check that:

```{r}
#get only patients with more than one result:
Aug_med_0vsABC.sum.7.2 <- Aug_med_0vsABC.sum.7 %>% 
  filter(num_of_results > 1)

#shuffle the data:
set.seed(1311)
Aug_med_0vsABC.sum.7.2 <- sample_frac(Aug_med_0vsABC.sum.7.2)  #1040 rows

#training and test sets:
n <- round(nrow(Aug_med_0vsABC.sum.7.2))
Aug_med_0vsABC.sum.7.2.train <- Aug_med_0vsABC.sum.7.2[1:(n*.75),]
Aug_med_0vsABC.sum.7.2.test <- Aug_med_0vsABC.sum.7.2[((n*.75)+1):n,]


randomForest(Result_2016 ~ min_result + max_result + var_result + 
             diff_of_last_and_max + diff_of_last_and_min, data=Aug_med_0vsABC.sum.7.2.train,
             xtest=select(Aug_med_0vsABC.sum.7.2.test, min_result, max_result, var_result, 
             diff_of_last_and_max, diff_of_last_and_min), ytest=Aug_med_0vsABC.sum.7.2.test$Result_2016,
             maxnodes = 200, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE)

137/140
```

Right. That's really good.  Let's try 21 days of data to see if we get any better results:

```{r}
#get only patients with more than one result:
Aug_med_0vsABC.sum.21.2 <- Aug_med_0vsABC.sum.21 %>% 
  filter(num_of_results > 1)

#shuffle the data:
set.seed(1312)
Aug_med_0vsABC.sum.21.2 <- sample_frac(Aug_med_0vsABC.sum.21.2)  #1040 rows

#training and test sets:
n <- round(nrow(Aug_med_0vsABC.sum.21.2))
Aug_med_0vsABC.sum.21.2.train <- Aug_med_0vsABC.sum.21.2[1:(n*.75),]
Aug_med_0vsABC.sum.21.2.test <- Aug_med_0vsABC.sum.21.2[((n*.75)+1):n,]


randomForest(Result_2016 ~ Age + Gender + BMI + AsaScore + TexturePancreas + min_result + max_result + var_result + 
             diff_of_last_and_max + diff_of_last_and_min, data=Aug_med_0vsABC.sum.21.2.train,
             xtest=select(Aug_med_0vsABC.sum.21.2.test, Age, Gender, BMI, AsaScore, TexturePancreas, min_result, max_result, var_result, 
             diff_of_last_and_max, diff_of_last_and_min), ytest=Aug_med_0vsABC.sum.21.2.test$Result_2016,
             maxnodes = 200, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE)

```

Not much different.

```{r}
#get only patients with more than one result:
Aug_med_AvsBvsC.sum.7.2 <- Aug_med_AvsBvsC.sum.7 %>% 
  filter(num_of_results > 1)

#shuffle the data:
set.seed(1312)
Aug_med_AvsBvsC.sum.7.2 <- sample_frac(Aug_med_AvsBvsC.sum.7.2) 

#training and test sets:
n <- round(nrow(Aug_med_AvsBvsC.sum.7.2))
Aug_med_AvsBvsC.sum.7.2.train <- Aug_med_AvsBvsC.sum.7.2[1:(n*.75),]
Aug_med_AvsBvsC.sum.7.2.test <- Aug_med_AvsBvsC.sum.7.2[((n*.75)+1):n,]


randomForest(Result_2016 ~ num_of_results + min_result + max_result + var_result + 
             diff_of_last_and_max + diff_of_last_and_min, data=Aug_med_AvsBvsC.sum.7.2.train,
             xtest=select(Aug_med_AvsBvsC.sum.7.2.test, num_of_results, min_result, max_result, var_result, 
             diff_of_last_and_max, diff_of_last_and_min), ytest=Aug_med_AvsBvsC.sum.7.2.test$Result_2016,
             maxnodes = 200, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE)

```

That is not good.  

```{r}
#get only patients with more than one result:
Aug_med_AvsBvsC.sum.21.2 <- Aug_med_AvsBvsC.sum.21 %>% 
  filter(num_of_results > 1)

#shuffle the data:
set.seed(1312)
Aug_med_AvsBvsC.sum.21.2 <- sample_frac(Aug_med_AvsBvsC.sum.21.2) 

#training and test sets:
n <- round(nrow(Aug_med_AvsBvsC.sum.21.2))
Aug_med_AvsBvsC.sum.21.2.train <- Aug_med_AvsBvsC.sum.21.2[1:(n*.75),]
Aug_med_AvsBvsC.sum.21.2.test <- Aug_med_AvsBvsC.sum.21.2[((n*.75)+1):n,]


randomForest(Result_2016 ~ num_of_results + min_result + max_result + var_result + 
             diff_of_last_and_max + diff_of_last_and_min, data=Aug_med_AvsBvsC.sum.21.2.train,
             xtest=select(Aug_med_AvsBvsC.sum.21.2.test, num_of_results, min_result, max_result, var_result, 
             diff_of_last_and_max, diff_of_last_and_min), ytest=Aug_med_AvsBvsC.sum.21.2.test$Result_2016,
             maxnodes = 200, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE)

```

No better with 3 weeks of data.

```{r}
randomForest(Result_2016 ~ Gender + Age + num_of_results + min_result + max_result + var_result + 
             diff_of_last_and_max + diff_of_last_and_min, data=Aug_med_AvsBvsC.sum.7.2.train,
             xtest=select(Aug_med_AvsBvsC.sum.7.2.test, Gender, Age, num_of_results, min_result, max_result, var_result, 
             diff_of_last_and_max, diff_of_last_and_min), ytest=Aug_med_AvsBvsC.sum.7.2.test$Result_2016,
             maxnodes = 200, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE, cutoff=c(0.45, 0.45, 0.1))
```

This might be suffering from an unbalanced data set.  Let's fix that first.

```{r}
table(Aug_med_AvsBvsC.sum.7.2$Result_2016)

set.seed(131)
Aug_med_AvsBvsC.sum.7.2.bal <- sample_frac(
  rbind(sample_n(filter(Aug_med_AvsBvsC.sum.7.2, Result_2016 == 'A'), 113),
        sample_n(filter(Aug_med_AvsBvsC.sum.7.2, Result_2016 == 'B'), 113),
        sample_n(filter(Aug_med_AvsBvsC.sum.7.2, Result_2016 == 'C'), 113)))

randomForest(Result_2016 ~ Gender + Age + BMI + AsaScore + TexturePancreas + min_result + max_result + var_result + 
             diff_of_last_and_max + diff_of_last_and_min, data=Aug_med_AvsBvsC.sum.7.2.bal,
             maxnodes = 200, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE)
             #cutoff=c(0.4, 0.4, 0.2))
```

This is still basically chance.  

```{r}
dim(Aug_med_AvsBvsC.sum.7.2)
sum(is.na(Aug_med_AvsBvsC.sum.7.2$OrganFailure))

#Too many missing values in:
#MPD_Diameter: 286 of 596
#Organ Failure: 221 of 596
#POD_DrainRemoval: 106 of 596

#Also Don't use:
#lastResult
#HospitalStay
#HospitalReadmission

#getting just the complete cases from with the columns above removed:
Aug_med_AvsBvsC.sum.7.2.cc <- Aug_med_AvsBvsC.sum.7.2[complete.cases(select(Aug_med_AvsBvsC.sum.7.2, ID, Gender, Age, BMI, AsaScore, TexturePancreas, min_result, max_result, var_result, diff_of_last_and_max, diff_of_last_and_min, Result_2016)),]

Aug_med_AvsBvsC.sum.7.2.cc <- Aug_med_AvsBvsC.sum.7.2.cc %>% 
  select(ID, Gender, Age, BMI, AsaScore, TexturePancreas, min_result, max_result, var_result, diff_of_last_and_max, diff_of_last_and_min, Result_2016)

table(Aug_med_AvsBvsC.sum.7.2.cc$Result_2016)

```

Let's now balance this dataset:

```{r}
set.seed(133)
Aug_med_AvsBvsC.sum.7.2.cc.bal <- sample_frac(
  rbind(sample_n(filter(Aug_med_AvsBvsC.sum.7.2.cc, Result_2016 == 'A'), 96),
        sample_n(filter(Aug_med_AvsBvsC.sum.7.2.cc, Result_2016 == 'B'), 96),
        sample_n(filter(Aug_med_AvsBvsC.sum.7.2.cc, Result_2016 == 'C'), 96)))

# how_many <- 70
# XA <- how_many/sum(Aug_med_AvsBvsC.sum.7.2.cc$Result_2016 == 'A')
# XB <- how_many/sum(Aug_med_AvsBvsC.sum.7.2.cc$Result_2016 == 'B')
# XC <- how_many/sum(Aug_med_AvsBvsC.sum.7.2.cc$Result_2016 == 'C')
# 
# Aug_med_AvsBvsC.sum.7.2.cc.bal <- Aug_med_AvsBvsC.sum.7.2.cc %>%
#   mutate(train_or_test = case_when(Result_2016 == 'A' ~ ifelse(runif() < XA, 0, 1),
#                                    Result_2016 == 'B' ~ ifelse(runif() < XB, 0, 1),
#                                    Result_2016 == 'C' ~ ifelse(runif() <  XC, 0, 1)))

Aug_med_AvsBvsC.sum.7.2.cc.bal
```

Let's get a training and test set and then randomForest everything!!!

```{r}
#training and test sets:
set.seed(133)
n <- round(nrow(Aug_med_AvsBvsC.sum.7.2.cc.bal))
Aug_med_ABC.cc.train <- Aug_med_AvsBvsC.sum.7.2.cc.bal[1:(n*.75),]
Aug_med_ABC.cc.test <- Aug_med_AvsBvsC.sum.7.2.cc.bal[((n*.75)+1):n,]

Aug_med_ABC.rf <- randomForest(Result_2016 ~ . - ID, data=Aug_med_ABC.cc.train,
             xtest=select(Aug_med_AvsBvsC.sum.7.2.cc, -Result_2016, -ID), ytest=Aug_med_AvsBvsC.sum.7.2.cc$Result_2016,
             maxnodes = 100, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE, 
             cutoff=c(0.36, 0.34, 0.3), mtry=8)

Aug_med_ABC.rf2 <- randomForest(Result_2016 ~ max_result + diff_of_last_and_max + diff_of_last_and_min, data=Aug_med_ABC.cc.train,
             xtest=select(Aug_med_AvsBvsC.sum.7.2.cc, max_result, diff_of_last_and_max, diff_of_last_and_min), ytest=Aug_med_AvsBvsC.sum.7.2.cc$Result_2016,
             maxnodes = 100, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE, 
             cutoff=c(0.36, 0.34, 0.3), mtry=3)


Aug_med_ABC.rf
```

```{r}
Aug_med_ABC.rf2
```

Let's get a balanced training set and then use everything else as the test set:

```{r}
set.seed(133)
Aug_med_AvsBvsC.sum.7.2.train2 <- sample_frac(
  rbind(sample_n(filter(Aug_med_AvsBvsC.sum.7.2.cc, Result_2016 == 'A'), 72),
        sample_n(filter(Aug_med_AvsBvsC.sum.7.2.cc, Result_2016 == 'B'), 72),
        sample_n(filter(Aug_med_AvsBvsC.sum.7.2.cc, Result_2016 == 'C'), 72)))
```

```{r}
Aug_med_AvsBvsC.sum.7.2.test2 <- filter(Aug_med_AvsBvsC.sum.7.2.cc, !ID %in% Aug_med_AvsBvsC.sum.7.2.train2$ID)
```

```{r}
Aug_med_ABC.rf <- randomForest(Result_2016 ~ . - ID, data=Aug_med_AvsBvsC.sum.7.2.train2,
             xtest=select(Aug_med_AvsBvsC.sum.7.2.test2, -Result_2016, -ID), ytest=Aug_med_AvsBvsC.sum.7.2.test2$Result_2016,
             maxnodes = 100, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE, 
             cutoff=c(0.36, 0.34, 0.3), mtry=6)

Aug_med_ABC.rf3 <- randomForest(Result_2016 ~ min_result + var_result + max_result + diff_of_last_and_max + diff_of_last_and_min, data=Aug_med_AvsBvsC.sum.7.2.train2,
             xtest=select(Aug_med_AvsBvsC.sum.7.2.test2, min_result, var_result, max_result, diff_of_last_and_max, diff_of_last_and_min), ytest=Aug_med_AvsBvsC.sum.7.2.test2$Result_2016,
             maxnodes = 100, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE, 
             cutoff=c(0.36, 0.34, 0.3), mtry = 3)

?randomForest
Aug_med_ABC.rf3
Aug_med_ABC.rf

library(pROC)

ABC.pred.rf <- predict(Aug_med_ABC.rf, Aug_med_AvsBvsC.sum.7.2.test2, type = 'prob')
ABC.mroc <- roc(ifelse(Aug_med_AvsBvsC.sum.7.2.test2$Result_2016 == 'C', 1, 0), ABC.pred.rf[,3], plot=T)
ABC.auc <- auc(ifelse(Aug_med_AvsBvsC.sum.7.2.test2$Result_2016 == 'C', 1, 0), ABC.pred.rf[,3])
coords(ABC.mroc, .4, "threshold", ret=c("sensitivity","specificity","ppv","npv"))

```

```{r}
Aug_med_test_prob <- as.tibble(predict(Aug_med_ABC.rf, Aug_med_ABC.cc.test, 'prob'))

Aug_med_test_prob$Result_2016 <- as.character(Aug_med_ABC.cc.test$Result_2016)

Aug_med_ABC.cc.train
```

Can I use `lime` to find the most important features?

```{r}
#in the online tutorial, they use caret::train.  Let's try that:
Aug_med_ABC.caret.rf <- caret::train(Result_2016 ~ .,
                         data = Aug_med_ABC.cc.train,
                         method = "rf")

Aug_med_ABC.caret.rf$finalModel
```

```{r}
explainer <- lime(Aug_med_ABC.cc.train, Aug_med_ABC.caret.rf, 
bin_continuous = FALSE, n_permutations = 100)

explanation <- lime::explain(Aug_med_ABC.cc.train[4,], explainer, n_labels=3, n_features=5)

windowsFonts(Times=windowsFont("TT Times New Roman"))
plot_features(explanation)


```

```{r}
Aug_med_train_prob <- predict(Aug_med_ABC.caret.rf, Aug_med_ABC.cc.train, 'prob') %>% 
  mutate(Result_2016=Aug_med_ABC.cc.train$Result_2016)

Aug_med_train_prob
```

```{r}
Aug_med_test_prob <- predict(Aug_med_ABC.caret.rf, Aug_med_ABC.cc.test, 'prob') %>% 
  mutate(Result_2016=Aug_med_ABC.cc.test$Result_2016)

Aug_med_test_prob

```

Let's make sure our 0 vs ABC random forest is okay with the added variables:

```{r}
Aug_med_0vsABC.sum.7.2.best_var <- Aug_med_0vsABC.sum.7.2 %>% 
  select(Gender, Age, BMI, AsaScore, TexturePancreas, min_result, max_result, var_result, diff_of_last_and_max, diff_of_last_and_min, Result_2016)

Aug_med_0vsABC.7.cc <- Aug_med_0vsABC.sum.7.2.best_var[complete.cases(Aug_med_0vsABC.sum.7.2.best_var),]

#shuffle the data:
set.seed(1311)
Aug_med_0vsABC.7.cc <- sample_frac(Aug_med_0vsABC.7.cc)  #834 rows

#training and test sets:
n <- round(nrow(Aug_med_0vsABC.7.cc))
Aug_med_0vsABC.cc.train <- Aug_med_0vsABC.7.cc[1:(n*.75),]
Aug_med_0vsABC.cc.test <- Aug_med_0vsABC.7.cc[((n*.75)+1):n,]


randomForest(Result_2016 ~ ., data=Aug_med_0vsABC.cc.train,
             xtest=select(Aug_med_0vsABC.cc.test, -Result_2016), ytest=Aug_med_0vsABC.cc.test$Result_2016,
             maxnodes = 200, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE)

Aug_med_0vsABC.cc.train
```

Let's now do A vs BC, as A vs B vs C was not working out so well.  I'll run this again, on Giulio's advice, with A vs BC:

```{r}
#get only patients with more than one result:
Aug_med_AvsBC.sum.7.2 <- Aug_med_AvsBC.sum.7 %>% 
  filter(num_of_results > 1)

#shuffle the data:
set.seed(1311)
Aug_med_AvsBC.sum.7.2 <- sample_frac(Aug_med_AvsBC.sum.7.2)  #596 rows

#using only complete cases with the variables we want:
Aug_med_AvsBC.cc <- Aug_med_AvsBC.sum.7.2 %>% 
  select(Gender, Age, BMI, AsaScore, TexturePancreas, min_result, max_result, var_result, diff_of_last_and_max, diff_of_last_and_min, Result_2016)

Aug_med_AvsBC.cc <- Aug_med_AvsBC.cc[complete.cases(Aug_med_AvsBC.cc),]

#balancing the data set:
#table(Aug_med_AvsBC.cc$Result_2016) A: 190, BC:301

set.seed(133)
Aug_med_AvsBC.cc.bal <- sample_frac(
  rbind(sample_n(filter(Aug_med_AvsBC.cc, Result_2016 == 'A'), 190),
        sample_n(filter(Aug_med_AvsBC.cc, Result_2016 == 'BC'), 190)))

table(Aug_med_AvsBC.cc.bal$Result_2016)
```

```{r}
#training and test sets:
n <- round(nrow(Aug_med_AvsBC.cc.bal))
Aug_med_AvsBC.train <- Aug_med_AvsBC.cc.bal[1:(n*.75),]
Aug_med_AvsBC.test <- Aug_med_AvsBC.cc.bal[((n*.75)+1):n,]

randomForest(Result_2016 ~ ., data=Aug_med_AvsBC.train,
             xtest=select(Aug_med_AvsBC.test, -Result_2016), ytest=Aug_med_AvsBC.test$Result_2016,
             maxnodes = 200, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE)
```

Well, 60% accuracy is not great.  Let's try AB vs C:


```{r}
#get only patients with more than one result:
Aug_med_ABvsC.sum.7.2 <- Aug_med_ABvsC.sum.7 %>% 
  filter(num_of_results > 1)

#shuffle the data:
set.seed(1311)
Aug_med_ABvsC.sum.7.2 <- sample_frac(Aug_med_ABvsC.sum.7.2)  #596 rows

#using only complete cases with the variables we want:
Aug_med_ABvsC.cc <- Aug_med_ABvsC.sum.7.2 %>% 
  select(Gender, Age, BMI, AsaScore, TexturePancreas, min_result, max_result, var_result, diff_of_last_and_max, diff_of_last_and_min, Result_2016)

Aug_med_ABvsC.cc <- Aug_med_ABvsC.cc[complete.cases(Aug_med_ABvsC.cc),]

#balancing the data set:
#table(Aug_med_AvsBC.cc$Result_2016) A: 190, BC:301

set.seed(133)
Aug_med_ABvsC.cc.bal <- sample_frac(
  rbind(sample_n(filter(Aug_med_ABvsC.cc, Result_2016 == 'AB'), 96),
        sample_n(filter(Aug_med_ABvsC.cc, Result_2016 == 'C'), 96)))

table(Aug_med_ABvsC.cc.bal$Result_2016)
```

```{r}
#training and test sets:
n <- round(nrow(Aug_med_ABvsC.cc.bal))
Aug_med_ABvsC.train <- Aug_med_ABvsC.cc.bal[1:(n*.75),]
Aug_med_ABvsC.test <- Aug_med_ABvsC.cc.bal[((n*.75)+1):n,]

randomForest(Result_2016 ~ ., data=Aug_med_ABvsC.train,
             xtest=select(Aug_med_ABvsC.test, -Result_2016), ytest=Aug_med_ABvsC.test$Result_2016,
             maxnodes = 100, type='prob', importance=TRUE, proximity = TRUE, keep.forest=TRUE)
```

Okay.  This is not going that well, but what is our goal.  We want a system that outputs a probability of being `C`.  It might be that the baseline probability is low.

```{r}
dim(filter(Aug_med, Result2016 == 'C'))[1]/length(Aug_med$Result2016)
```

The baseline, without knowing anything about the patient is 9.7%.

Once we separate them into 0 vs ABC, the probability goes up to:

```{r}
dim(filter(Aug_med, Result2016 == 'C'))[1]/dim(filter(Aug_med, Result2016 %in% c('A', 'B', 'C')))[1]
```

Now the probability is 21.37%.

We just need a way to modify that probability.  Not too much, but how do we do that??  When in doubt, visualize:

```{r, fig.height=7, fig.width=10}
ggplot(Aug_med_AvsBvsC.sum.7.2.cc, aes(x=as.factor(Age), y=BMI, color=Result_2016)) +
  geom_point()
```

```{r, fig.height=7, fig.width=10}
ggplot(filter(Aug_med_AvsBvsC.sum.7.2.cc, var_result < 1000000), aes(x=Result_2016, y=diff_of_last_and_max+1, fill=as.factor(cut_number(BMI,4)))) +
  geom_boxplot() +
  scale_y_log10()

```

```{r, fig.height=7, fig.width=10}
ggplot(Aug_med_AvsBvsC.sum.7.2.cc, aes(x=as.factor(Gender), y=BMI, color=Result_2016)) +
  geom_jitter(height = 0)
```

```{r, fig.height=7, fig.width=10}
ggplot(Aug_med_AvsBvsC.sum.7.2.cc, aes(x=as.factor(Gender), y=TexturePancreas, color=Result_2016)) +
  geom_jitter(height = 0.3)
```

```{r, fig.height=7, fig.width=10}
ggplot(filter(Aug_med_AvsBvsC.sum.7.2.cc, diff_of_last_and_max < 100000), aes(x=as.factor(Gender), y=(diff_of_last_and_max+1), color=Result_2016)) +
  geom_jitter(height = 0.1) +
  scale_y_log10()
```

```{r, fig.height=7, fig.width=10}
ggplot(filter(Aug_med_AvsBvsC.sum.7.2.cc, diff_of_last_and_max < 100000), aes(x=var_result+1, y=(diff_of_last_and_max+1), color=Result_2016)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10()

```

```{r}

```
